---
title: "simpleRegress"
author: "Kimberly Aldana "
date: "Fall 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```


The scatter plot function is used to visualize the relationship between the "NECK" and "LENGTH" variables, with "NECK" values represented on the x-axis and "LENGTH" values represented on the y-axis. The abline function adds a line to the plot, which represents the best-fit line determined by the linear regression model. The bear's neck is 20 inches around which helps predict about 57 in length. Looking at the accompanying data, the residual standard error is 5.376 and the R-squared is .7524.






## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 


```{r}
MYCHEST <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(MYCHEST)
summary.lm(MYCHEST)
```

The plot function creates a scatter plot where the "CHEST" values are plotted on the x-axis and the corresponding "LENGTH" values are plotted on the y-axis. The abline function adds a line to the plot, which represents the best-fit line determined by the linear regression model.The residual standard error is 4.939 and its lower than the previous model which is 5.376 better. In addition, the R-squared is 0.791 which is higher than the previous (0.791) and also better. As of now, ranking between chest and neck, chest would be the clear winner on top.



```{r}
MYHEADWTH <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(MYHEADWTH)
summary.lm(MYHEADWTH)
```

The plot function generates a scatter plot where the "HEADWTH" values are plotted on the x-axis and the corresponding "LENGTH" values are plotted on the y-axis. The abline function adds a line to the plot that represents the best-fit line determined by the linear regression model. As the third model, the residual error is 7.32 which is making it worst. In addition, the R-squared is now 0.5409 which is also a decline from the previous model. Comparing all three models, NECKWTH would be the bottom of the list. The residual error and R-squared numbers show clear low numbers. This would rank #3 of all 3.






