---
title: "simpleRegress"
author: "Kimberly Aldana "
date: "Fall 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```


The scatter plot function is used to visualize the relationship between the "NECK" and "LENGTH" variables, with "NECK" values represented on the x-axis and "LENGTH" values represented on the y-axis. The abline function adds a line to the plot, which represents the best-fit line determined by the linear regression model. The bear's neck is 20 inches around which helps predict about 57 in length. Looking at the accompanying data, the residual standard error is 5.376 and the R-squared is .7524.






## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 


```{r}
MYCHEST <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(MYCHEST)
summary.lm(MYCHEST)
```

The plot function creates a scatter plot where the "CHEST" values are plotted on the x-axis and the corresponding "LENGTH" values are plotted on the y-axis. The abline function adds a line to the plot, which represents the best-fit line determined by the linear regression model.The residual standard error is 4.939 and its lower than the previous model which is 5.376 better. In addition, the R-squared is 0.791 which is higher than the previous (0.791) and also better. As of now, ranking between chest and neck, chest would be the clear winner on top.



```{r}
MYHEADWTH <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(MYHEADWTH)
summary.lm(MYHEADWTH)
```

The plot function generates a scatter plot where the "HEADWTH" values are plotted on the x-axis and the corresponding "LENGTH" values are plotted on the y-axis. The abline function adds a line to the plot that represents the best-fit line determined by the linear regression model. As the third model, the residual error is 7.32 which is making it worst. In addition, the R-squared is now 0.5409 which is also a decline from the previous model. Comparing all three models, NECKWTH would be the bottom of the list. The residual error and R-squared numbers show clear low numbers. This would rank #3 of all 3.


```{r}
MYWEIGHT <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(MYWEIGHT)
summary.lm(MYWEIGHT)
```

This model is in representation of the bear's weight. The residual standard error is 5.432 and the adjusted R-squared is 0.7472. In this case it is higher than the last model which is worse, and the adjusted R-square is higher than the previous model, which is better. So whereas, the residual standard error is worse, the adjusted R-square is better. At this point the bear's chest has the highest adjusted R-square which is the best. As far as the ranking goes, out of all the graphs we examined, WEIGHT would be considered 3rd place while HEADWTH takes the bottom of the list.



```{r}
MYAGE <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(MYAGE)
summary.lm(MYAGE)
```

After close examination of the graph, the residual standard error comes up to be about 7.511, while the adjusted R-squared comes out to be 0.5073. With the R-squared number to be extremely low, it comes out to be the worst R-squared we have come across. Whilst looking at the residual standard error, 7.511 is a high number, which makes it being the worst. By far, AGE is clicking on the bottom of the list while CHEST comes at 1st.



```{r}
MYHEADLEN <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(MYHEADLEN)
summary.lm(MYHEADLEN)
```

While giving this graph a close examination, the residual standard error is 4.235, while the adjusted R-squared is 0.8434. With this record breaking numbers by far, HEADLEN takes first place! With the residual standard error number being the lowest, and R-squared being the highest, we can conclude that this takes the number 1 spot out of all the graphs.


#CONCLUSION 

After placing multiple graphs to conclude which factor had the winning crown, I have concluded that MYHEADLEN takes first place! It contributed to a low residual standard error of 4.235 and a high R-squared number of 0.8434, it takes the number 1 spot out of all the other factors. Taking second place is CHEST, taking third place is NECK, taking fourth place is WEIGHT, taking fifth place is HEADWTH and taking last of all is AGE, being the worst one by far. 


