---
title: "simpleRegress"
author: "Kimberly Aldana "
date: "Fall 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```


The scatter plot function is used to visualize the relationship between the "NECK" and "LENGTH" variables, with "NECK" values represented on the x-axis and "LENGTH" values represented on the y-axis. The abline function adds a line to the plot, which represents the best-fit line determined by the linear regression model. The bear's neck is 20 inches around which helps predict about 57 in length. Looking at the accompanying data, the residual standard error is 5.376 and the R-squared is .7524.






## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 


```{r}
MYCHEST <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(MYCHEST)
summary.lm(MYCHEST)
```









